---
title: "00_data_input_and_manipulation"
author: "Bella Oleksy"
date: "4/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Document history
I created this R markdown file on April 12, 2019 after submitting the manuscript to Ecosystems on April 1, 2019.
Based off of "SkyLochMaster_R_NEWANDIMPROVED.R" which was created November 29, 2018 after correcting GAMs and visuals with advice from Jared Wolfe

LVDC3_Master_RawR_170329.csv is the version I used for a long time.
Now switching to LVDC3_Master_RawR_180208.csv ... only difference is
that I added a column for date, which has the interpolation from Jack Cornett's original spreadsheet
very approximate, still waiting on dates from Peter.

####Load packages
```{r }

library(tidyverse)
library(ggpubr)
library(ggplot2)
library(grid) #for function unit() in ggplot2
library(gtable) #for adding subtitle in ggplot2
library(ggthemes)
library(scales) # to add percentages to plots and pretty_breaks()
library(gridExtra) #for arranging graphs side by side w ftn grid.arrange()
library(skimr)#Summary statistics
library(reshape2)

#load packages necessary for GAMS
library("mgcv")
library("scam")
library("cowplot")
library("grid")                         # for unit.pmax(), unit.list()
library("schoenberg")
library("tidyr")
library("nlme")


```

##Plot themes
```{r}
library(ggthemes)
library(ggplot2)
library(ggpmisc)
library(ggpubr)
#Customize theme 
#http://joeystanley.com/blog/custom-themes-in-ggplot2 
theme_paleo <- function () { 
  theme_base(base_size=8, base_family="Helvetica") %+replace% 
    theme(
      panel.background  = element_blank(),
      plot.background = element_rect(fill="white", colour=NA),
      plot.title=element_text(face="plain",hjust=0),
      plot.subtitle = element_text(color="dimgrey", hjust=0, size=8),
      axis.title=element_text(face="plain"),
      panel.grid.minor=element_blank(),
      strip.text.x = element_text(face="italic")
    )
}
# , panel.grid.major=element_line(color="lightgrey")
theme_set(theme_paleo())  # pre-set the bw theme
```


#SPDC1 Data Cleaning 

```{r, echo=FALSE}
SPDC1_raw <- read.csv("data/raw/SPDC1_RawR_171205_nmolorgC.csv")

SPDC1_raw <- na.omit(SPDC1_raw)
tibble::as_tibble(SPDC1_raw)
#Converts data to tbl class. tbl’s are easier to examine than data frames.
#R displays only the data that fits onscreen

glimpse(SPDC1_raw) #quick view of the data
# View(SPDC1_raw) #looks the same as before-- good, didn't fuck it up already

#condence tbl to only include variables of interest
#then create new variable chla_pheoa which adds chlorophyll a and pheo a together
SPDC1_short <- SPDC1_raw %>%
  mutate(chla_pheoa = chl_a + pheo_a,
         chlb_pheob = chl_b + phaeo_b,
         all_chl = chl_a + pheo_a + chl_ap,
         all_diat = diato + diadino,
         PI = chl_a/pheo_a,
         UVR_index = sed_C/(allo+diato+lutein),
         AD_ratio = allo/diato,
         master_chl = (chl_a + chl_b)/(chl_ap+pheo_a+phaeo_b),
         perc_chlb = (phaeo_b/pheo_a),
         perc_diatox = (diato/chla_pheoa)) %>%
  select(sample_start_depth, depth_plot, depth_range, sample_ID, lake_ID:del15N, fuco, oscil, aphan, diadino, myxo, allo, diato,
         lutein, cantha, b_car, chl_ap, echine, pheo_a, phaeo_b, chla_pheoa, chlb_pheob,
         PI, UVR_index, all_chl, all_diat, AD_ratio,master_chl,perc_chlb,perc_diatox)
glimpse(SPDC1_short) #it worked!!
str(SPDC1_short)

#calculate z-scores of d15N
SPDC1_short$del15N_zscore<-scale(SPDC1_short$del15N, center = TRUE, scale = TRUE)




```

##Depth-age function
```{r}
# a simple function assuming a constant sedimentation rate, but in reality this
# function will probably use the approx()
# function given known age/depth values from 210Pb or other dating method.
# http://apps.fishandwhistle.net/archives/1287
depth_to_age_sky <- function(depth, sed_rate = 0.059, core_year = 2017) {
  # sed_rate here is in cm/year
  core_year - 1 / sed_rate * depth
}
SPDC1_short$age_sky<-depth_to_age_sky(SPDC1_short$depth_plot)
#Sedimentation rate from Leavitt lab = 0.059 cm/year

#Make a dataframe of just ages and depths for Anna
sky_export <- SPDC1_short %>%
  select(sample_start_depth:lake_ID, age_sky, percN)
write_csv(sky_export, "data/output/sky_depths_dates.csv")
```

##Wide to long
```{r}
#Convert SPDC1_short from wide to long
SPDC1_long = melt(SPDC1_short, id.vars = c("lake_ID", "age_sky", "sample_ID", "sample_start_depth"),
                  measure.vars = c("percC", "percN",
                                   "CN", "del13C", "del15N", "fuco","oscil","aphan",
                                   "diadino","myxo","allo","diato","lutein","cantha",
                                   "b_car","chl_ap","echine","pheo_a","phaeo_b","chla_pheoa","chlb_pheob",
                                   "PI","UVR_index","all_chl","all_diat","AD_ratio","master_chl","perc_chlb","perc_diatox","del15N_zscore")) %>%
  arrange(desc(age_sky)) %>%
  dplyr::rename(proxy_ID = variable, 
         response = value)
```

## Diatoms
```{r}

taxon_category <- read.csv("data/raw/taxon_category.csv")
diatoms_long_raw <- read.csv("data/raw/SPDC1_180808_diatomsquant_IAO_long.csv")

diatoms_long_PB <- left_join(diatoms_long_raw,taxon_category, by="spp_ID")
# write.csv(diatoms_long_PB, file="diatom_BP.csv")
glimpse(diatoms_long_PB)

diatoms_wide_PB <- dcast(diatoms_long_PB, INSTAAR_accession_num + PLANK_B ~ spp_ID, value.var="counts")

diatoms_wide_withSPP <- diatoms_wide_PB %>%
  mutate(total = select(., Achnanthidium:Tetracyclus) %>%
           rowSums(na.rm=TRUE)) 

diatoms_wide_PB <- diatoms_wide_withSPP %>%
  select(INSTAAR_accession_num, PLANK_B, total)

diatoms_wide_PB <- dcast(diatoms_wide_PB, INSTAAR_accession_num ~ PLANK_B, value.var="total") %>%
  select(INSTAAR_accession_num, PLANK, B) %>%
  mutate(BP = B/PLANK,
         PB = PLANK/B) %>%
  select(-PLANK, -B)

diatom_metadata <- read.csv("/Users/solomonlab/Google Drive/Research (common)/Research/Data/Sediment Cores/SPDC1/R/diatom_metadata.csv")

diatom_full <- left_join(diatom_metadata, diatoms_wide_PB, by="INSTAAR_accession_num")


depth_to_age_sky_diatoms <- function(depth, sed_rate = 0.059, core_year = 2017) {
  # sed_rate here is in cm/year
  core_year - 1 / sed_rate * depth
}
diatom_full$age_sky<-depth_to_age_sky_diatoms(diatom_full$depth_plot)
#Sedimentation rate from Leavitt lab = 0.059 cm/year
```

#### 2019-11-27 Explore diatoms...
```{r}
df1 <- diatom_full %>%
  select(INSTAAR_accession_num, age_sky)

specieswithdate<-left_join(diatoms_wide_withSPP, df1, by=c("INSTAAR_accession_num"))

specieswithdate <- specieswithdate %>%
  mutate(Stauroneis_spp = sum(Stauroneis + Staurosira + Staurosirella, na.rm=TRUE),
         Karayevia_spp = sum(Karayevia + `Karayevia clevei`, na.rm=TRUE),
         Fragilaria_spp = sum(Fragilaria + `Fragilaria crotonensis` + Fragilariforma, na.rm=TRUE)) %>%
  select(-`None detected`)
# write.csv(specieswithdate, file="data/output/specieswithdates.csv")
### Had trouble getting the rows to sum so had to do a little manual excel data summary. Ugh.
specieswithdate <- read.csv("/Users/solomonlab/Google Drive/Research (common)/Research/Data/R/paleo/data/output/specieswithdates.csv")

B_P_separate<- specieswithdate %>%
  filter(age_sky > 1850) %>%
ggplot(aes(x=age_sky,y=total, shape=PLANK_B, color=PLANK_B))+
  # annotate("rect", xmin=1850, xmax=2020, ymin=0, ymax=Inf, alpha=0.1, fill="grey10") +
  geom_point(size=4) +
  # geom_line(size=1)+
  # geom_ribbon(aes(ymin = (lower), ymax = (upper), x = age_sky, group=lake_ID), alpha = 0.2, inherit.aes = FALSE, fill = "black") +
  coord_cartesian(xlim=c(1850,2010))+
  coord_cartesian(ylim=c(0,100))+
  # scale_y_continuous(breaks=seq(0, 20, 4))+
  scale_x_continuous(breaks=seq(1850, 2010, 50))+
  scale_color_manual(
    name="",
    breaks=c("B","PLANK"),
    labels=c("Benthic","Planktonic"),
    values=c("#5e4fa2", "#f46d43"))+
  scale_shape_manual(
    name="",
    breaks=c("B","PLANK"),
    labels=c("Benthic","Planktonic"),
    values = c(17, 19))+
  labs(y="Total # valves counted", x="\nYear")+
  theme(legend.position="bottom", legend.text = element_text(size=28))+
  guides(shape = guide_legend(override.aes = list(size=10)),
         color = guide_legend(override.aes = list(linetype = 0)))
B_P_separate

specieswithdate2 <- specieswithdate %>%
  select(-c("Stauroneis", "Staurosira","Staurosirella",
          "Nitzschia", "Nitzschia.dissipata",
          "Karayevia","Karayevia.clevei",
          "Fragilaria", "Fragilaria.crotonensis",
          "Fragilariforma", "total"))
specieswithdatelong = melt(specieswithdate2,
                           id.vars = c("INSTAAR_accession_num", "age_sky","PLANK_B"),
                           measure.vars = c("Achnanthidium", "Asterionella.formosa",
                                            "Aulacoseira", "Cavinula", "Chamaepinnularia", "Cymbopleura",
                                            "Encyonema", "Eolimna", "Eunotia",
                                            "Genkalia","Gomphonema",
                                            "Hantzschia", "Humidophila", 
                                            "Luticola", "Mayamaea",
                                            "Melosira.varians", "Navicula",
                                            "Nupela", "Pinnularia", "Platessa",
                                            "Psammothidium", "Rossithidium",
                                            "Sellaphora", "Tetracyclus",
                                            "Stauroneis_spp", "Karayevia_spp",
                                            "Fragilaria_spp", "Nitzschia_spp" )) %>%
  arrange(variable)


# specieswithdatelong<-left_join(diatoms_long_PB, df1, by=c("INSTAAR_accession_num"))

# unique(specieswithdatelong$variable)
library(RColorBrewer)
nb.cols <- 28
mycolors <- colorRampPalette(brewer.pal(8, "Accent"))(nb.cols)
# Create a ggplot with 18 colors 
# Use scale_fill_manual
ggplot(df) + 
  geom_col(aes(name, Sepal.Length, fill = factor(Sepal.Length))) +
  scale_fill_manual(values = mycolors) +
  theme_minimal() +
  theme(legend.position = "top")

```

###** MS supplemental figure
```{r}

specieswithdate_wide <- specieswithdate %>%
  select("Achnanthidium", "Asterionella.formosa",
            "Aulacoseira", "Cavinula", "Chamaepinnularia", "Cymbopleura",
            "Encyonema", "Eolimna", 
            "Genkalia", "Humidophila", 
             "Pinnularia", 
            "Psammothidium", 
            "Sellaphora", 
            "Stauroneis_spp", "Karayevia_spp",
            "Fragilaria_spp", "Nitzschia_spp",
         "INSTAAR_accession_num",
         "age_sky",
         "PLANK_B") %>%
  rename("Asterionella formosa"="Asterionella.formosa",
         "Stauroneis"="Stauroneis_spp",
         "Karayevia"="Karayevia_spp",
         "Fragilaria crotonensis"="Fragilaria_spp",
         "Nitzschia"="Nitzschia_spp")


specieswithdatelong = melt(specieswithdate_wide,
                           id.vars = c("INSTAAR_accession_num", "age_sky","PLANK_B"),
                           measure.vars = c("Achnanthidium", "Asterionella formosa",
                                            "Aulacoseira", "Cavinula", "Chamaepinnularia", "Cymbopleura",
                                            "Encyonema", "Eolimna",
                                            "Genkalia", "Humidophila", 
                                             "Pinnularia", 
                                            "Sellaphora",
                                            "Stauroneis", "Karayevia",
                                            "Fragilaria crotonensis", "Nitzschia" ))



specieswithdatalong_trim <- specieswithdatelong %>%
  filter(PLANK_B != "NA")%>%
  # mutate(is.numeric(as.integer()))%>%
  filter(age_sky > 1850) %>%
  mutate(value = na_if(value, "0"))
str(specieswithdatalong_trim)



specieswithdatalong_trim$variable <- factor(specieswithdatalong_trim$variable,
                                 levels=c("Asterionella formosa",
                                           "Aulacoseira",
                                          "Fragilaria crotonensis",
                                     "Stauroneis",
                                          "Achnanthidium",
                                          "Cavinula",
                                          "Chamaepinnularia",
                                          "Cymbopleura",
                                           "Encyonema", "Eolimna",
                                           "Genkalia", "Humidophila", 
                                            "Pinnularia", 
                                           "Sellaphora",
                                          "Karayevia",
                                            "Nitzschia"))
#More of a time series
specieswithdatalong_trim%>%
ggplot(aes(x=age_sky,y=value, color=PLANK_B))+
  geom_point(size=2.5, alpha=0.8) + 
    coord_cartesian(xlim=c(1850,2020))+
  scale_x_continuous(breaks=seq(1850, 2020, 50))+
  scale_y_continuous(breaks=pretty_breaks(n=5))+
  scale_color_manual(
    name="",
    breaks=c("B","PLANK"),
    labels=c("Benthic","Planktonic"),
    values=c("#5e4fa2", "#f46d43"))+
  labs(y="Total # valves counted", x="Year")+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle = 45, vjust=0.55))+
  guides(shape = guide_legend(override.aes = list(size=50)),
         color = guide_legend(override.aes = list(linetype = 0)))+
  facet_wrap(variable~., ncol=4)




#More like traditional paleo analyses

paleo_diatoms<-specieswithdatalong_trim %>%
  ggplot(aes(y = age_sky, x = value, color=PLANK_B )) +
  geom_segment(aes(xend = 0, yend = age_sky), lwd = 1) +
  facet_grid(.~variable, scales = "free_x", space = "free_x") +
  scale_x_continuous(breaks = c(0, 25, 50, 75)) +
  # scale_y_reverse() +
  scale_color_manual(
    name="",
    breaks=c("B","PLANK"),
    labels=c("Benthic","Planktonic"),
    values=c("#5e4fa2", "#f46d43")) +
  labs(x = "# valves counted", y = "Year") +
  theme(strip.text.x = element_text(angle = 60, hjust = 0, vjust = 0), 
        strip.background = element_blank(),
        legend.position="bottom")

paleo_diatoms

# voodoo that makes it so that facet labels can overlap
# https://stackoverflow.com/questions/49740215/ggplot-facet-grid-label-cut-off
species_plot_grob <- ggplotGrob(paleo_diatoms)
for(i in which(grepl("strip-t", species_plot_grob$layout$name))){
  species_plot_grob$grobs[[i]]$layout$clip <- "off"
}

# needed to draw the modified plot_grob
grid::grid.draw(species_plot_grob)

ggsave("figures/MS/FigureSX.diatom_valve_counts_resubmission.jpg", width=18, height=18,units="in")

```



###** MS  figure - rel abundance
```{r}

specieswithdate_rel_abund_wide <- specieswithdate %>%
  select("Achnanthidium", "Asterionella.formosa",
            "Aulacoseira", "Cavinula", "Chamaepinnularia", "Cymbopleura",
            "Encyonema", "Eolimna", 
            "Genkalia", "Humidophila", 
             "Pinnularia", 
            "Psammothidium", 
            "Sellaphora", 
            "Stauroneis_spp", "Karayevia_spp",
            "Fragilaria_spp", "Nitzschia_spp",
         "INSTAAR_accession_num",
         "age_sky",
         "PLANK_B",
         "total") %>%
  rename("Asterionella formosa"="Asterionella.formosa",
         "Stauroneis"="Stauroneis_spp",
         "Karayevia"="Karayevia_spp",
         "Fragilaria crotonensis"="Fragilaria_spp",
         "Nitzschia"="Nitzschia_spp")



specieswithdatelong_relabund = melt(specieswithdate_rel_abund_wide,
                           id.vars = c("INSTAAR_accession_num", "age_sky","PLANK_B","total"),
                           measure.vars = c("Achnanthidium", "Asterionella formosa",
                                            "Aulacoseira", "Cavinula", "Chamaepinnularia", "Cymbopleura",
                                            "Encyonema", "Eolimna",
                                            "Genkalia", "Humidophila", 
                                             "Pinnularia", 
                                            "Sellaphora",
                                            "Stauroneis", "Karayevia",
                                            "Fragilaria crotonensis", "Nitzschia" )) %>%
    filter(PLANK_B != "NA")%>%
    filter(age_sky > 1850) %>%
  mutate(value = na_if(value, "0"),
         relAbund=(value/total)*100)
str(specieswithdatelong_relabund)



specieswithdatelong_relabund$variable <- factor(specieswithdatelong_relabund$variable,
                                 levels=c("Asterionella formosa",
                                           "Aulacoseira",
                                          "Fragilaria crotonensis",
                                     "Stauroneis",
                                          "Achnanthidium",
                                          "Cavinula",
                                          "Chamaepinnularia",
                                          "Cymbopleura",
                                           "Encyonema", "Eolimna",
                                           "Genkalia", "Humidophila", 
                                            "Pinnularia", 
                                           "Sellaphora",
                                          "Karayevia",
                                            "Nitzschia"))


#More like traditional paleo analyses

#Function for making my own pretty breaks
breaks_fun <- function(x) {
  if (max(x) < 20) {
    seq(0, 20, 4)
  } else {
    seq(0, 100, 25)
  }
}

paleo_diatoms<-specieswithdatelong_relabund %>%
  ggplot(aes(y = age_sky, x = relAbund, color=PLANK_B )) +
  geom_segment(aes(xend = 0, yend = age_sky), lwd = 1) +
  facet_grid(.~variable, scales="free_x", space = "free_x") +
  scale_y_continuous(breaks = c(1850,1875,1900,1925,1950,1975,2000)) +
  scale_x_continuous(breaks=breaks_fun) +
  scale_color_manual(
    name="",
    breaks=c("B","PLANK"),
    labels=c("Benthic","Planktonic"),
    values=c("#5e4fa2", "#f46d43")) +
  labs(x = "Relative Abundance (%)", y = "Year") +
  theme(strip.text.x = element_text(angle = 60, hjust = 0, vjust = 0), 
        strip.background = element_blank(),
        legend.position="bottom",
        plot.margin=unit(c(0,1,1,0), "lines"),
        # panel.spacing=grid::unit(0,"lines")
        )

paleo_diatoms

# # voodoo that makes it so that facet labels can overlap
# # https://stackoverflow.com/questions/49740215/ggplot-facet-grid-label-cut-off
# species_plot_grob <- ggplotGrob(paleo_diatoms)
# for(i in which(grepl("strip-t", species_plot_grob$layout$name))){
#   species_plot_grob$grobs[[i]]$layout$clip <- "off"
# }
# 
# # needed to draw the modified plot_grob
# grid::grid.draw(species_plot_grob)
# 
# ggsave("figures/MS/FigureS5.diatom_relabundance_resubmission.jpg", width=8, height=5,units="in")


png(
  "figures/MS/FigureS5.diatom_relabundance_resubmission.png", 
  width = 9, 
  height = 3, 
  res = 300,
  units = "in"
)
species_plot_grob <- ggplotGrob(paleo_diatoms)
for(i in which(grepl("strip-t", species_plot_grob$layout$name))){
  species_plot_grob$grobs[[i]]$layout$clip <- "off"
}

# needed to draw the modified plot_grob
grid::grid.draw(species_plot_grob)
dev.off()
```

#LVDC3 Data Cleaning
```{r}
LVDC3_raw <- read.csv("data/raw/LVDC3_Master_RawR_180208.csv")

LVDC3_raw <- na.omit(LVDC3_raw)
as_tibble(LVDC3_raw)

LVDC3_short <- LVDC3_raw %>%
  mutate(chla_pheoa = chl_a + pheo_a,
         chlb_pheob = chl_a + phaeo_b,
         PI = chl_a/pheo_a,
         UVR_index = sed_C/(allo+diato+lutein),
         all_diat = diadino + diato,
         AD_ratio = allo/diato,
         all_chl = chl_ap + chl_a + pheo_a,
         master_chl = (chl_a + chl_b)/(chl_ap + phaeo_b + pheo_a),
         perc_chlb = (phaeo_b/pheo_a),
         perc_diatox = (diato/chla_pheoa)) %>%
  select(depth_plot, depth_range, sample_ID, 
         lake_ID, depth_plot:fuco, oscil, aphan, diadino, myxo, allo, diato, lutein, b_car, cantha, echine,
         chl_ap, chl_b, chl_a, phaeo_b, pheo_a, PI, UVR_index, chla_pheoa, chlb_pheob,
         all_diat, all_chl, percN, percC, master_chl, AD_ratio,perc_chlb,perc_diatox)
glimpse(LVDC3_short) #it worked!!

LVDC3_short$del15N_zscore<-scale(LVDC3_short$del15N, center = TRUE, scale = TRUE)

```

##Depth-age function
```{r}

depth_to_age_loch <- function(depth, sed_rate = 0.059, core_year = 2016) {
  # sed_rate here is in cm/year
  core_year - 1 / sed_rate * depth
}
LVDC3_short$age_sky<-depth_to_age_loch(LVDC3_short$depth_plot)

#Make a dataframe of just ages and depths for 
loch_export <- LVDC3_short %>%
  select(depth_plot:lake_ID, age_sky, percN)
write_csv(loch_export, "data/output/loch_depths_dates.csv")
```


##Wide to long
```{r}
#Convert LVDC3_short from wide to long
LVDC3_long = melt(LVDC3_short, id.vars = c("lake_ID", "age_sky", "sample_ID", "depth_plot"),
                  measure.vars = c("percC", "percN",
                                   "CN", "del13C", "del15N", "fuco","oscil","aphan",
                                   "diadino","myxo","allo","diato","lutein","cantha",
                                   "b_car","chl_ap","echine","pheo_a","phaeo_b","chla_pheoa","chlb_pheob",
                                   "PI","UVR_index","all_chl","all_diat","AD_ratio","master_chl","perc_chlb","perc_diatox","del15N_zscore")) %>%
  arrange(desc(age_sky)) %>%
  dplyr::rename(proxy_ID = variable, #rename columns 
         response = value,
         sample_start_depth = depth_plot)
str(LVDC3_long)
str(SPDC1_long)

```



#Combine Loch & Sky to Data
```{r}

#Combine both Loch and Sky into one master dataset
LVDC3_long_merge <- LVDC3_long %>%
  select("lake_ID","age_sky","sample_start_depth","proxy_ID","response")
SPDC1_long_merge <- SPDC1_long %>%
  select("lake_ID","age_sky","sample_start_depth","proxy_ID","response")

master_core_long <- bind_rows(LVDC3_long_merge, SPDC1_long_merge)

#Specify lake_ID as factor
master_core_long$lake_ID <- as.factor(master_core_long$lake_ID)

#Convert master from long to wide
master_core_wide <- dcast(master_core_long, lake_ID + age_sky + sample_start_depth ~ proxy_ID, value.var="response") 

loch_wide <- master_core_wide %>%
  filter(lake_ID=="TheLoch")

sky_wide <- master_core_wide %>%
  filter(lake_ID=="SkyPond")
```


#EXPORT

```{r}

write.csv(master_core_wide, file="data/output/master_core_wide.csv")
write.csv(master_core_long, file="data/output/master_core_long.csv")
write.csv(loch_wide, file="data/output/loch_wide.csv")
write.csv(sky_wide, file="data/output/sky_wide")
write.csv(diatom_full, "data/output/diatom_full.csv")

```